---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
library(tidyverse)

deaths <- av |> 
  pivot_longer(
    cols = starts_with('Death'),
    names_to = 'Time',
    values_to = 'Died'
    )
deaths$Time <- parse_number(deaths$Time)

returns <- av |> 
  pivot_longer(
    cols = starts_with('Return'),
    names_to = 'Return Number',
    values_to = 'Returned'
  )
returns$`Return Number` <- parse_number(returns$`Return Number`)

deaths |> count(Died)

deaths$URL |> unique() |> length()

89/173
```
An Avengers suffers an average number of 0.51445 deaths

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### Teammate name: Colin Huntley
### FiveThirtyEight Statement

> "I counted 89 total deaths."

### Include the code

```{r}
sum(deaths$Died == "YES")
```

### Include your answer

According to the reults, there are 89 total deaths, which backs up what the original analysis found.


### Teammate name: Nate Riedl
### FiveThirtyEight Statement

> "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team."

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
deaths |> count(Time, Died)
```

### Include your answer

According to the results of the code I ran, there are 69 Avengers who died at least one time, which matches the number found on the website from the above question.


### Teammate name: Prudvik Seemakurthi
### FiveThirtyEight Statement

> "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife."

### Code:

```{r}
# Filter for first deaths and first returns
first_deaths <- deaths %>%
  filter(Time == 1, Died == "YES")

first_returns <- returns %>%
  filter(`Return Number` == 1, Returned == "YES")

# Merge the two filtered sets using URL
merged <- first_deaths %>%
  left_join(first_returns, by = "URL", suffix = c("_death", "_return"))

# Compute totals and probability
summary <- first_deaths %>%
  summarise(
    total_died = n(),
    total_returned = sum(URL %in% first_returns$URL),
    prob_returned = total_returned / total_died
  )

summary

```


### Answer:

After checking the Avengers data, I found that about two out of three heroes who died came back to life after their first death. This means the statement there’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife is true based on the data.


### Teammate name: Selena Cooper
### FiveThirtyEight Statement

> "but only a 50 percent chance they recovered from a second or third death."

### Include the code
```{r}
recover2 <- deaths |> 
  group_by(Name.Alias) |> 
  summarize(
    death_count = sum(Died == "YES"),
    return_count = sum(Returned == "YES")
  ) |>
  filter(death_count >= 2) |> 
  summarize(
    total_case = n(),
    recovered_from_2nd_or_3rd = sum(return_count >= 2),
    no_recovery = total_case - recovered_from_2nd_or_3rd,
    percent_recovered_2nd_or_3rd = recovered_from_2nd_or_3rd / total_case
  )

recover2
```

### Include your answer

The claim states that only 50% of characters recover from a second or third death. Using the dataset, 45 out of 64 characters who died at least once recovered from a second or third death, giving an actual recovery rate of 70.3%. Therefore, the claim underestimates the true recovery rate, and is disproven by the data.
